---
title: "Pepsi Cola Project"
author: "James"
date: "2025-02-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The original dataset contained data for both Coca-Cola and Pepsi-Cola in a single sheet. To enhance clarity and facilitate analysis, the data was split into two separate sheets: ‘brand1’ for Coca-Cola and ‘brand2’ for Pepsi-Cola. Both sheets retain the same structure, ensuring consistency in analysis.

## Question 1
Calculate descriptive (summary) statistics about the sales and use of marketing mix variables for both brands. Who is the market leader? How do the brands compare in terms of pricing, promotion, assortment? How do the brands compare in terms of allocation of spending between the four advertising instruments? Report the statistics and discuss your learnings and insights.

```{r message=FALSE, warning=FALSE}
# Load necessary libraries
library(readxl)  # For reading Excel files
library(dplyr)   # For data manipulation
library(knitr)   # For displaying tables

# Read data from the Excel file
brand1 <- read_excel("MA_assignment_data.xls", sheet = "brand1")
brand2 <- read_excel("MA_assignment_data.xls", sheet = "brand2")

# Summary statistics for Coca-Cola (brand1) excluding 'week' column
summary_brand1 <- brand1 %>%
  select(-week) %>%  # Exclude the 'week' column
  summarise(across(where(is.numeric), list(
    Min = ~min(. , na.rm = TRUE), 
    Max = ~max(. , na.rm = TRUE), 
    Mean = ~mean(. , na.rm = TRUE), 
    Median = ~median(. , na.rm = TRUE), 
    SD = ~sd(. , na.rm = TRUE)
  )))

# Transpose the summary table for Coca-Cola
summary_brand1_transposed <- t(summary_brand1)

# Summary statistics for Pepsi-Cola (brand2) excluding 'week' column
summary_brand2 <- brand2 %>%
  select(-week) %>%  # Exclude the 'week' column
  summarise(across(where(is.numeric), list(
    Min = ~min(. , na.rm = TRUE), 
    Max = ~max(. , na.rm = TRUE), 
    Mean = ~mean(. , na.rm = TRUE), 
    Median = ~median(. , na.rm = TRUE), 
    SD = ~sd(. , na.rm = TRUE)
  )))

# Transpose the summary table for Pepsi-Cola
summary_brand2_transposed <- t(summary_brand2)

# Display results in Markdown format
cat("\n### Summary Statistics for Coca-Cola (Brand 1):\n")
kable(summary_brand1_transposed)

cat("\n### Summary Statistics for Pepsi-Cola (Brand 2):\n")
kable(summary_brand2_transposed)





```

### Market Leadership:

Based on the summary statistics, **Coca-Cola** appears to be the market leader in terms of sales volume. The total weekly sales for Coca-Cola (brand1) have a **mean of 308,246 liters** and a **maximum of 476,880 liters**, while Pepsi-Cola (brand2) has a **mean of 44,901 liters** and a **maximum of 84,319 liters**. This significant difference in sales suggests Coca-Cola is the dominant brand in the market.

### Pricing:

In terms of pricing, Coca-Cola has a slightly higher **average price** per liter at **€1.349** compared to Pepsi-Cola's **€1.045**. Coca-Cola's price range also spans from **€1.27** to **€1.42**, while Pepsi-Cola's price range is lower, ranging from **€0.83** to **€1.13**. This indicates that Coca-Cola is priced higher, potentially reflecting its premium market positioning.

### Promotion:

For promotions, Coca-Cola has a significantly higher **average display percentage (14.26%)** compared to Pepsi-Cola’s **4.51%**, which suggests that Coca-Cola has a stronger presence on the shelf in terms of visible displays. Additionally, Coca-Cola has a **higher average feature percentage (4.11%)** compared to Pepsi-Cola’s **4.37%**. Despite this, Coca-Cola’s stronger display presence suggests a more aggressive promotional strategy.

### Assortment:

Coca-Cola also has a slightly larger **average assortment** (44.83 SKUs) compared to Pepsi-Cola’s **12.96 SKUs**. This indicates that Coca-Cola offers a broader range of products, which could appeal to a wider variety of consumer preferences, giving it an advantage in terms of product availability and choice.

### Advertising Spending Allocation:

When comparing advertising spend, Coca-Cola leads in most categories:
- **TV advertising**: Coca-Cola spends an average of **€378,137** per week, significantly higher than Pepsi-Cola's **€58,190**. This suggests Coca-Cola is more aggressive in traditional media advertising.
- **Digital advertising**: Coca-Cola also has higher digital ad spending (**€4,816**) compared to Pepsi-Cola’s **€288**.
- **OOH advertising**: Coca-Cola spends **€82,106** on Out-Of-Home advertising, while Pepsi-Cola only spends **€7,687**.
- **Magazine advertising**: Coca-Cola’s magazine ad spending averages **€13,701**, compared to Pepsi-Cola’s **€782**.

Overall, Coca-Cola allocates a higher proportion of its marketing budget across all four advertising instruments, with a particularly large focus on TV and OOH advertising.

### Conclusion:

Coca-Cola is the clear market leader, outpacing Pepsi-Cola in sales volume, pricing, promotional activity, and advertising spend. The higher sales, larger assortment, more extensive promotional activity, and greater investment in advertising solidify Coca-Cola's dominance in the market. Pepsi-Cola, while competitive, appears to have a more restrained marketing strategy, focusing less on promotion and advertising.

```{r message=FALSE, warning=FALSE}
# Load necessary library
library(tidyverse)

# Assuming the dataset has already been loaded as brand2 data frame
# Add 1 only to variables that contain zero values for feature, display, tv, digital, ooh, and magazine
brand2 <- brand2 %>%
  mutate(
    log_sales = log(sales.brand2),
    log_feature = ifelse(feature.brand2 == 0, log(feature.brand2 + 1), log(feature.brand2)),
    log_display = ifelse(display.brand2 == 0, log(display.brand2 + 1), log(display.brand2)),
    log_assortment = log(assortment.brand2),
    log_price = log(price.brand2),
    log_tv = ifelse(tv.brand2 == 0, log(tv.brand2 + 1), log(tv.brand2)),
    log_digital = ifelse(digital.brand2 == 0, log(digital.brand2 + 1), log(digital.brand2)),
    log_ooh = ifelse(ooh.brand2 == 0, log(ooh.brand2 + 1), log(ooh.brand2)),
    log_magazine = ifelse(magazine.brand2 == 0, log(magazine.brand2 + 1), log(magazine.brand2))
  )

# Estimate the regression model
model <- lm(log_sales ~ log_feature + log_display + log_assortment + log_price + 
              log_tv + log_digital + log_ooh + log_magazine, data = brand2)

# Display the summary of the model
summary(model)

```
### Analysis of Log-Log Regression Model Results for Pepsi:

In this analysis, we estimated the log-log regression model for Pepsi-Cola’s sales, with log-transformed marketing mix variables. For the variables where zero values existed (feature, display, TV, digital, OOH, and magazine), we added a value of **1** to ensure the log transformation could be applied. This adjustment is standard practice when dealing with variables that contain zero values.

The estimated model is:

\[
\log(\text{sales.brand2}) = \beta_0 + \beta_1 \log(\text{feature.brand2} + 1) + \beta_2 \log(\text{display.brand2} + 1) + \beta_3 \log(\text{assortment.brand2}) + \beta_4 \log(\text{price.brand2}) + \beta_5 \log(\text{tv.brand2} + 1) + \beta_6 \log(\text{digital.brand2} + 1) + \beta_7 \log(\text{ooh.brand2} + 1) + \beta_8 \log(\text{magazine.brand2} + 1)
\]

#### Model Overview:
- **Multiple R-squared**: **0.6412**, indicating that about **64.12%** of the variation in Pepsi’s sales is explained by the marketing mix variables.
- **Adjusted R-squared**: **0.6128**, confirming a reasonably good model fit.
- **F-statistic**: **22.56** with a **p-value < 2.2e-16**, suggesting that the model is statistically significant.

#### Coefficients and Interpretation:

1. **Intercept (β0)**: **10.04** (statistically significant), representing the baseline sales when all variables are at their baseline levels (after log transformation).

2. **log_feature (β1)**: **-0.7038** (statistically significant, p < 0.001), indicating that higher feature support (after adding 1) is associated with lower sales.

3. **log_display (β2)**: **-0.0055** (not statistically significant, p = 0.429), suggesting display support does not have a substantial effect on sales.

4. **log_assortment (β3)**: **0.2456** (not statistically significant, p = 0.188), indicating that the range of products offered by Pepsi does not significantly affect sales in this dataset.

5. **log_price (β4)**: **-0.3370** (not statistically significant, p = 0.268), suggesting that price changes are not significantly related to sales.

6. **log_tv (β5)**: **-0.0015** (not statistically significant, p = 0.521), showing no significant relationship between TV ad spending and sales.

7. **log_digital (β6)**: **0.0004** (not statistically significant, p = 0.937), suggesting digital ad spending has negligible impact on sales.

8. **log_ooh (β7)**: **0.0053** (not statistically significant, p = 0.123), indicating out-of-home advertising does not have a significant effect on sales.

9. **log_magazine (β8)**: **0.0109** (not statistically significant, p = 0.268), implying magazine advertising does not significantly contribute to sales.

#### Conclusion:
- The **only significant variable** is **log_feature**, indicating that Pepsi's feature support has an inverse relationship with sales, suggesting a decrease in sales as feature support increases.
- **Other marketing mix variables**, including display, assortment, price, TV, digital, OOH, and magazine, do not show statistically significant relationships with Pepsi's sales in this model.

This analysis highlights the importance of feature support in Pepsi’s sales, while other marketing activities may need further examination for their effectiveness.




